Cooling Requirements for the Vertical Shear Instability in Protoplanetary Disks

### 2015-04-19 paper.tex
minor edits to youdin's new intro, updated notations and reference styles

To reach hydrodynamic turbulence via the VSI, a disk should  
first develop the linear instability. A quantitative 
examination of the thermodynamic requirement of the linear VSI is thus essential in assessing its 
relevance to PPDs. In this work, we study the linear VSI in vertically global disk models 
with non-zero cooling timescales, which permit buoyancy forces. 
Our effort generalizes both early analyses of the VSI that account for buoyancy but in vertically localized disk models \citep{urpin98,urpin03}, 
as well as recent calculations in vertically global disks but without buoyancy forces \citepalias{nelson13,mcnally14,barker15}. 
%to study of how finite cooling times, and thus buoyancy, 
%affect the VSI. Our effort may also be regarded
%as a extension of early analyses of the VSI in 
%vertically localized disk models with cooling \citep{urpin98,urpin03} to
%vertically global disks with cooling. 
We consider disk   
models with vertically constant $t_{\rm c}$, as well as more realistic
cooling timescales based on dust opacity in both the optically thick
(diffusive) and thin regimes, where appropriate. Coupled with the
results from numerical simulations, our results can be used to 
understand how the VSI operates over a wide range of model PPDs.

>>>>

To reach hydrodynamic turbulence via the VSI, a disk should  
first develop the linear instability. A quantitative 
examination of the thermodynamic requirement of the linear VSI is thus essential in assessing its 
relevance to PPDs. In this work, we study the linear VSI in vertically global disk models 
with non-zero cooling timescales, which permit buoyancy forces. 
Our effort generalizes both early analyses of the VSI that account for buoyancy but in vertically localized disk models \citep{urpin98,urpin03}, and
recent calculations in vertically global disks but without buoyancy forces \citepalias{nelson13,mcnally14,barker15}. 
%to study of how finite cooling times, and thus buoyancy, 
%affect the VSI. Our effort may also be regarded
%as a extension of early analyses of the VSI in 
%vertically localized disk models with cooling \citep{urpin98,urpin03} to
%vertically global disks with cooling. 
We consider disk   
models with vertically constant $t_{\rm c}$, as well as more realistic
cooling timescales based on dust opacity in both the optically thick
(diffusive) and thin regimes, where appropriate. Coupled with the
results from numerical simulations, our results can be used to 
understand how the VSI operates over a wide range of model PPDs.
